---
title: "Medicaid recertification requirements study - Prepare study data"
output: html_document
---

```{r}

# Clear space

rm(list=ls())

# Load packages

library(dhsDW)
library(ROracle)
library(tidyverse)

# Set chunk options

knitr::opts_chunk$set(eval = FALSE)

options(scipen=999)

# Set file path

filepath <- 'C:/Users/K011014/OneDrive - Allegheny County/Medicaid/Effect of Medicaid recertification deadlines project/Data analysis'
```

# Create the base panel data sets of Medicaid enrollment relative to recert deadlines

## Pull base files

### Medicaid enrollment spells for all-time, among adults age 18 to 64 who were enrolled in MAGI Medicaid at some point since 4/1/2023

```{r}

#-------------------------------------------------------------------------
# Pull raw data

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

sample_enrollment_dates_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)

select distinct
  a.ma_receipt_num, 
  a.mci_uniq_id,
  c.elig_from_date, 
  c.elig_to_date
from 
  medicaid_magi_adults_04012023 a
  left join ac_mh.ccbh_hc_eligibility c on c.ma_receipt_num = a.ma_receipt_num')

# Make var names lowercase

names(sample_enrollment_dates_raw) <- tolower(names(sample_enrollment_dates_raw))

#-------------------------------------------------------------------------
# Clean up the data

# Reformat dates as R date format

sample_enrollment_dates <- sample_enrollment_dates_raw %>%
  mutate_at(vars(elig_from_date, elig_to_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d')) %>%
  
# Remove any enrollment spells where the start date is after the end date. These are data quality issues that I think we can safely remove. 
  
  filter(elig_from_date < elig_to_date) %>%
  
# Recode end dates of 6/6/2079 as missing. 
  
  mutate(elig_to_date = if_else(elig_to_date == as.Date('2079-06-06'), NA, elig_to_date)) %>%
  
# Deduplicate the data by enrollment spell
  
  distinct()
  
  # Save data

save(sample_enrollment_dates, file = paste0(filepath, "/Data/Medicaid enrollment data/sample_enrollment_dates.rda"))
```

### Medicaid recertification deadlines between 4/1/2023 and today, among adults age 18 to 64 who were enrolled in MAGI Medicaid at some point since 4/1/2023

```{r}

#-------------------------------------------------------------------------
#-------------------------------------------------------------------------
# Pull raw data

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

sample_recert_deadlines_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)
    
select distinct
    a.ma_receipt_num, 
    a.mci_uniq_id,
    c.ma_renewal_date as annual_recert_deadline, 
    c.ma_sar_date as semiannual_recert_deadline
from 
    medicaid_magi_adults_04012023 a 
    inner join ac_mh.pa_client_elig_daily c on c.recipient_nbr = a.ma_receipt_num
where
    c.ma_renewal_date is not null 
    or c.ma_sar_date is not null')

# Make var names lowercase

names(sample_recert_deadlines_raw) <- tolower(names(sample_recert_deadlines_raw))

#-------------------------------------------------------------------------
#-------------------------------------------------------------------------
# Clean up the data

# Convert dates to R date format

sample_recert_deadlines <- sample_recert_deadlines_raw %>%
  mutate_at(vars(semiannual_recert_deadline, annual_recert_deadline), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d')) %>%
  
# Reshape the data to have one row per deadline
  
  gather(variable, recert_deadline, -ma_receipt_num, -mci_uniq_id) %>%
  select(-variable) %>%
  
# Deduplicate the data
  
  filter(!is.na(recert_deadline)) %>%
  distinct(ma_receipt_num, mci_uniq_id, recert_deadline) %>%
  
# Keep only the recertification deadlines from between 4/1/2023 and 6/30/2025 
  
  filter(recert_deadline >= as.Date('2023-04-01') & recert_deadline <= as.Date('2025-06-30'))

#-------------------------------------------------------------------------
# Remove any recertification deadlines that were >=2 deadlines into the future at the time the person left Medicaid - we can safely remove these deadlines b/c they were non-binding and they could not have induced a pre-treatment voluntary withdrawal. For example, let’s say a person enrolls in Medicaid on 1/1/2024. They are assigned a 6-month recert deadline of 6/1/2024 and a 12-month deadline of 1/1/2025. They voluntarily withdraw from Medicaid on 5/25/2024 and do not re-enroll before 1/1/2025. In this case we’d want to keep the 6/1/2024 deadline in the data but drop the 1/1/2025 deadline. The 1/1/2025 deadline plays no role in the person’s decision-making, and if we retain it in the data then it only serves to dilute the treatment effects because it’s an empty, meaningless deadline.

# Load the Medicaid enrollment data for the sample

load(file = paste0(filepath, "/Data/Medicaid enrollment data/sample_enrollment_dates.rda"))

# Identify the recertification deadlines that are at least two deadlines after the last deadline on which the person was enrolled in Medicaid

sample_recert_deadlines <- sample_recert_deadlines %>%
  left_join(sample_enrollment_dates, by = c('ma_receipt_num', 'mci_uniq_id')) %>%
  mutate(elig_to_date = replace_na(elig_to_date, Sys.Date()),
         on_medicaid_as_of_deadline = ifelse(recert_deadline >= elig_from_date & recert_deadline <= elig_to_date, 1, 0)) %>%
  group_by(ma_receipt_num, mci_uniq_id, recert_deadline) %>%
  summarise(on_medicaid_as_of_deadline = ifelse(sum(on_medicaid_as_of_deadline) > 0, 1, 0)) %>%
  mutate(prior_recert_deadline = dplyr::lag(recert_deadline)) %>%
  ungroup() %>%
  arrange(ma_receipt_num, mci_uniq_id, recert_deadline) %>%
  group_by(ma_receipt_num, mci_uniq_id) %>%
  mutate(multiple_deadlines_since_being_on_medicaid = ifelse(cumsum(on_medicaid_as_of_deadline) == dplyr::lag(cumsum(on_medicaid_as_of_deadline), n = 2L), 1, 0)) %>%
  ungroup() %>%
  mutate(multiple_deadlines_since_being_on_medicaid = replace_na(multiple_deadlines_since_being_on_medicaid, 0)) %>%

# Now see if the person was enrolled in Medicaid at any point between the flagged "non-binding" deadline and the most recent prior binding deadline. If not, then we can drop the flagged deadline. If so, then keep it because it was still binding in the sense that it was the person's known upcoming deadline.

  left_join(sample_enrollment_dates, by = c('ma_receipt_num', 'mci_uniq_id')) %>%
  mutate(drop = ifelse(elig_to_date > prior_recert_deadline & elig_to_date < recert_deadline, 1, 0)) %>%
  group_by(ma_receipt_num, mci_uniq_id, recert_deadline, multiple_deadlines_since_being_on_medicaid) %>%
  summarise(drop = ifelse(sum(drop, na.rm = TRUE) > 0, 1, 0)) %>%
  ungroup() %>%
  filter(multiple_deadlines_since_being_on_medicaid + drop < 2) %>%
  select(ma_receipt_num, mci_uniq_id, recert_deadline)
  
# Save data

save(sample_recert_deadlines, file = paste0(filepath, "/Data/Medicaid enrollment data/sample_recert_deadlines.rda"))
```

### Medicaid termination dates and reasons between 4/1/2023 and today, among adults age 18 to 64 who were enrolled in MAGI Medicaid at some point since 4/1/2023

```{r}

#-------------------------------------------------------------------------
# Pull raw data

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

sample_terminations_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)

select distinct
    a.*,
    c.batch_date as termination_date, 
    c.ma_term_reason_code as termination_reason_code,
    e.ma_term_reason_desc as termination_reason_desc
from 
    medicaid_magi_adults_04012023 a 
    left join ac_mh.pa_client_elig_daily c on c.recipient_nbr = a.ma_receipt_num 
    left join ac_mh.LKP_PA_MA_TERM_REASON e on e.ma_term_reason_code = c.ma_term_reason_code
where 
    c.change_type = \'Terminated Coverage\'')

# Make var names lowercase

names(sample_terminations_raw) <- tolower(names(sample_terminations_raw))

#---------------------------------------------------------------------------
# Clean up data

# Reformat dates as R date format

sample_terminations <- sample_terminations_raw %>%
  mutate(termination_date = as.Date(substring(as.character(termination_date), 1, 10), '%Y-%m-%d'))

# Save data

save(sample_terminations, file = paste0(filepath, "/Data/Medicaid enrollment data/sample_terminations.rda"))
```

## Create panel data

```{r}

#----------------------------------------------------------------------------
#----------------------------------------------------------------------------
#----------------------------------------------------------------------------
# Load relevant data

# Medicaid enrollment dates among the base sample (all adults who were enrolled in MAGI Medicaid at some point since 4/1/2023)

load(file = paste0(filepath, "/Data/Medicaid enrollment data/sample_enrollment_dates.rda"))

# Medicaid recert deadlines since 4/1/2023 among the base sample

load(file = paste0(filepath, "/Data/Medicaid enrollment data/sample_recert_deadlines.rda"))

#----------------------------------------------------------------------------
#----------------------------------------------------------------------------
#----------------------------------------------------------------------------
# Create panel data

#----------------------------------------------------------------------------
#----------------------------------------------------------------------------
# Day-level panel

sample_recert_deadlines <- sample_recert_deadlines %>%
    slice_sample(n = 10000) 

#----------------------------------------------------------------------------
# Create a grid of each day within +/- 365 days of each of the person's recert deadlines. Keep only the recert deadlines that give us at least 365 days of outcome data for every person in the base sample. For now, this is the recert deadlines up to 6/30/2024, since our enrollment data only goes through 6/30/2025 at the moment. We can extend this on a rolling basis. 

# Filter to desired deadline date range

recert_date_window_365 <- sample_recert_deadlines %>%
  filter(recert_deadline <= as.Date('2024-06-30')) %>%
  
# Create a unique ID for each person and recert deadline

  arrange(mci_uniq_id, recert_deadline) %>%
  group_by(mci_uniq_id) %>%
  mutate(deadline_counter = sequence(n())) %>%
  ungroup() %>%
  mutate(person_deadline_id = paste0(mci_uniq_id, "_", deadline_counter)) %>%
  
# Create the grid of dates

  mutate(start_date = recert_deadline - 365, 
         end_date = pmin(recert_deadline + 365, as.Date('2025-06-30'))) %>%
  group_by(ma_receipt_num, mci_uniq_id, person_deadline_id, recert_deadline) %>%
  reframe(date = seq.Date(from = start_date,
                         to = end_date,
                         by = "day")) %>%
  ungroup() %>%
  mutate(tau_days = as.numeric(date - recert_deadline), 
         
# Create a variable for calendar date fixed-effects

         calendar_day = as.numeric(floor(difftime(date, min(date), units = "days"))) + 1)

#----------------------------------------------------------------------------
# Create a panel with one obs per person per day that they were enrolled in Medicaid between 4/1/2022 and 6/30/2025

# Remove enrollment episodes where the start date is after 6/30/2025

enrollment_days <- sample_enrollment_dates %>%
  inner_join(sample_recert_deadlines %>% distinct(ma_receipt_num), by = c('ma_receipt_num')) %>%
  filter(elig_from_date <= as.Date('2025-06-30')) %>%
  
# Recode missing values of the enrollment episode end date as 6/30/2025
  
  mutate(elig_to_date = replace_na(elig_to_date, as.Date('2025-06-30'))) %>%
  
# Remove enrollment episodes that ended before 4/1/2022 (one year before the earliest recert date). We don't need to look back any further than this for now.
  
  filter(elig_to_date >= as.Date('2022-04-01')) %>%
  
# Recode any episode start dates as 4/1/2022 if they are before 4/1/2022
  
  mutate(elig_from_date = if_else(elig_from_date < as.Date('2022-04-01'), as.Date('2022-04-01'), elig_from_date)) %>%
  
# Deduplicate the data
  
  distinct(ma_receipt_num, mci_uniq_id, elig_from_date, elig_to_date) %>%
  
# Reshape the data to have 1 obs for each day during each enrollment episode
  
  group_by(ma_receipt_num, mci_uniq_id, elig_from_date, elig_to_date) %>%
  reframe(date = list(seq.Date(from = elig_from_date,
                               to = elig_to_date,
                               by = "day"))) %>%
  unnest(date) %>%
  ungroup() %>%
  distinct(ma_receipt_num, mci_uniq_id, date) %>%
  mutate(medicaid_enrolled = 1) 

#----------------------------------------------------------------------------
# Combine the enrollment days with the grid of dates around each recert deadline

medicaid_enrollment_panel_daily <- enrollment_days %>%
  right_join(recert_date_window_365, by = c('ma_receipt_num', 'mci_uniq_id', 'date')) %>%
  mutate(medicaid_enrolled = replace_na(medicaid_enrolled, 0),
  
# Create a treatment indicator 
  
         treat = ifelse(tau_days >= 0, 1, 0)) 
  
# Save data
  
save(medicaid_enrollment_panel_daily, file = paste0(filepath, "/Data/Medicaid enrollment data/medicaid_enrollment_panel_daily.rda"))

#----------------------------------------------------------------------------
#----------------------------------------------------------------------------
# Week-level panel

# Roll up the daily panel into weeks

medicaid_enrollment_panel_weekly <- medicaid_enrollment_panel_daily %>%
  mutate(tau_weeks = as.numeric(floor(difftime(date, recert_deadline, units = "weeks"))) + 1,
         calendar_week = as.numeric(floor(difftime(date, min(date), units = "weeks"))) + 1) %>%
  group_by(ma_receipt_num, mci_uniq_id, person_deadline_id, recert_deadline, tau_weeks, calendar_week) %>%
  summarise(medicaid_enrolled = ifelse(sum(medicaid_enrolled) > 0, 1, 0)) %>%
  ungroup() %>%
  mutate(treat = ifelse(tau_weeks >= 0, 1, 0))
  
# Save data
  
save(medicaid_enrollment_panel_weekly, file = paste0(filepath, "/Data/Medicaid enrollment data/medicaid_enrollment_panel_weekly.rda"))
```

# Create outcomes data sets

## Clinical Connect alerts data

### Client data coverage dates

```{r}

#-------------------------------------------------------------------------
# Pull raw data

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

clinical_connect_coverage_dates <- dbGetQuery(conn = db_con, 
                          statement = 
'select 
   a.mci_uniq_id,
   b.first_date,
   change_date as latest_date
from ac_hcdm.cchie_client_master a
inner join
(select a.mci_uniq_id,
 min(a.load_Date) as first_date
from ac_hcdm.cchie_client_master_hist a
group by mci_uniq_id) b
on a.mci_uniq_id = b.mci_uniq_id')

# Make var names lowercase

names(clinical_connect_coverage_dates) <- tolower(names(clinical_connect_coverage_dates))

#------------------------------------------------------------------------------
# Clean the data

# Reformat dates as R date format

clinical_connect_coverage_dates <- clinical_connect_coverage_dates %>%
  mutate_at(vars(first_date, latest_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(clinical_connect_coverage_dates, file = paste0(filepath, "/Data/Clinical Connect health care data/clinical_connect_coverage_dates.rda"))
```

## Hospital event data

```{r}

#-------------------------------------------------------------------------
# Pull raw data

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

clinical_connect_events_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)

select distinct
    a.*,
    e.encounter_id,
    e.event_type, 
    e.patient_class, 
    e.admission_type, 
    e.event_date_time,
    e.admit_date_time,
    e.discharge_date_time,
    e.admit_source,
    e.admission_reason,
    e.diagnosis_code_1,
    e.diagnosis_description_1, 
    e.diagnosis_code_2,
    e.diagnosis_description_2, 
    e.diagnosis_code_3,
    e.diagnosis_description_3, 
    e.diagnosis_code_4,
    e.diagnosis_description_4, 
    e.diagnosis_code_5,
    e.diagnosis_description_5
from 
    medicaid_magi_adults_04012023 a
    inner join 
        (select distinct
            mci_uniq_id, 
            uniqueid
        from 
            AC_HCDM.CCHIE_CLIENT_MASTER) c on c.mci_uniq_id = a.mci_uniq_id 
    inner join ac_hcdm.cchie_alert e on LPAD(e.roster_patient_id, 10, \'0\') = c.uniqueid')

# Make var names lowercase

names(clinical_connect_events_raw) <- tolower(names(clinical_connect_events_raw))

#------------------------------------------------------------------------------
# Clean the data

# Reformat dates as R date format

clinical_connect_events <- clinical_connect_events_raw %>%
  mutate_at(vars(event_date_time, admit_date_time, discharge_date_time), ~ as.Date(substring(as.character(.), 1, 10), '%Y%m%d'))

# Save data

save(clinical_connect_events, file = paste0(filepath, "/Data/Clinical Connect health care data/clinical_connect_events.rda"))
```

## Medical examiner data

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

medical_examiner_deaths <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct
    a.*, 
    e.caldr_dt,
    c.ME_ALL_ACT_IND,
    c.ME_ACC_OD_ALL_ACT_IND,
    c.ME_ACC_OD_OPD_ACT_IND,
    c.ME_ACC_OD_NON_OPD_ACT_IND,
    c.ME_ACC_OD_UNK_ACT_IND,
    c.ME_HOM_ACT_IND,
    c.ME_SUIC_ACT_IND,
    c.ME_OTH_ACT_IND
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    inner join dw.fact_client_prog_status_month @DW c on c.mci_uniq_id = a.mci_uniq_id 
    left join dw.dim_time @DW e on e.time_key = c.time_key
where
    c.ME_ALL_ACT_IND + c.ME_ACC_OD_ALL_ACT_IND + c.ME_ACC_OD_OPD_ACT_IND + c.ME_ACC_OD_NON_OPD_ACT_IND + 
    c.ME_ACC_OD_UNK_ACT_IND + c.ME_HOM_ACT_IND + c.ME_SUIC_ACT_IND + c.ME_OTH_ACT_IND > 0
    and e.caldr_dt >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(medical_examiner_deaths) <- tolower(names(medical_examiner_deaths))

# Reformat dates to R date format

medical_examiner_deaths <- medical_examiner_deaths %>%
  mutate_at(vars(caldr_dt), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d')) 

# Save data

save(medical_examiner_deaths, file = paste0(filepath, "/Data/Medicaid claims data/medical_examiner_deaths.rda"))
```

## UPMC claims data since 1/1/2022

```{r, eval = TRUE}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

upmc_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select 
    a.*,
    c.CLAIM_NBR, 
    c.claim_line_nbr, 
    c.service_begin_date,
    c.ICD_PROCEDURE_ID,
    c.icd_procedure_desc, 
    c.cpt4_code, 
    c.cpt4_desc,
    e.icd_diagnosis_id, 
    e.icd_diagnosis_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    inner join ac_upmc.upmc_claim_proc c on c.mci_uniq_id = a.mci_uniq_id
    left join ac_upmc.upmc_claim_dx e on e.master_member_id = c.master_member_id and e.claim_nbr = c.claim_nbr
where
    c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(upmc_claims) <- tolower(names(upmc_claims))

# Reformat dates to R date format

upmc_claims <- upmc_claims %>%
  mutate_at(vars(service_begin_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(upmc_claims, file = paste0(filepath, "/Data/UPMC claims data/upmc_claims.rda"))
```

## Medicaid claims data

### Physical health claims since 1/1/2022

```{r, eval = TRUE}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

#--------------------------------------------------------------------------
# Non-ER inpatient care

non_er_inpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_INPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Non-Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(non_er_inpatient_claims) <- tolower(names(non_er_inpatient_claims))

# Reformat dates to R date format

non_er_inpatient_claims <- non_er_inpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(non_er_inpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/non_er_inpatient_claims.rda"))

#--------------------------------------------------------------------------
# ER inpatient care

er_inpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_INPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(er_inpatient_claims) <- tolower(names(er_inpatient_claims))

# Reformat dates to R date format

er_inpatient_claims <- er_inpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(er_inpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/er_inpatient_claims.rda"))

#--------------------------------------------------------------------------
# Non-ER outpatient care

non_er_outpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_OUTPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Non-Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(non_er_outpatient_claims) <- tolower(names(non_er_outpatient_claims))

# Reformat dates as R date format

non_er_outpatient_claims <- non_er_outpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(non_er_outpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/non_er_outpatient_claims.rda"))

#--------------------------------------------------------------------------
# ER outpatient care

er_outpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_OUTPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(er_outpatient_claims) <- tolower(names(er_outpatient_claims))

# Reformat dates as R date format

er_outpatient_claims <- er_outpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(er_outpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/er_outpatient_claims.rda"))
```

### Mental health claims since 1/1/2022

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

#---------------------------------------------------------------------------------
# Substance use disorder service claims

sud_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct
    a.*,
    c.claim_number, 
    c.svc_start_dt, 
    c.svc_end_dt, 
    c.payer, 
    c.office, 
    c.service_code, 
    c.service_name, 
    c.cost_ctr_common_desc, 
    c.service_category, 
    c.prvdr_name, 
    c.cnty_tot, 
    c.tot_units, 
    c.diagnosis_desc, 
    c.diagnosis_common_sort
from (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
inner join dare_health.sud_fy1518 @DW c on a.mci_uniq_id = c.mci_uniq_id
where
    c.svc_start_dt >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(sud_claims) <- tolower(names(sud_claims))

# Reformat dates in R date format

sud_claims <- sud_claims %>%
    mutate_at(vars(svc_start_dt, svc_end_dt), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(sud_claims, file = paste0(filepath, "/Data/Medicaid claims data/sud_claims.rda"))

#---------------------------------------------------------------------------------
# Non-SUD behavioral health claims

mh_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct
    a.*,
    c.service_rendered_row_id, 
    c.svc_start_dt, 
    c.svc_end_dt, 
    c.payer,
    c.office, 
    c.service_code,
    replace(c.service_name, chr(0), \' \') as service_name,
    c.cost_center, 
    c.service_category, 
    c.provider_name,
    c.cnty_tot, 
    c.tot_units, 
    c.diagnosis_code,
    c.diagnosis_desc, 
    c.diagnosis_common_sort, 
    c.diagnosis_cat
from  (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
inner join dare_health.mh_svc_tableau @DW c on a.mci_uniq_id = c.mci_uniq_id
where
    c.svc_start_dt >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(mh_claims) <- tolower(names(mh_claims))

# Reformat dates in R date format

mh_claims <- mh_claims %>%
    mutate_at(vars(svc_start_dt, svc_end_dt), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(mh_claims, file = paste0(filepath, "/Data/Medicaid claims data/mh_claims.rda"))

#---------------------------------------------------------------------------------
# Payments from the Allegheny County behavioral health managed care organization to health care providers

# Note that these amounts are not the cost of care to Medicaid. This is just an alternative way of quantifying the volume of a person's health care utilization. The 'total charges' field is the amount that the provider billed to CCBH. The 'prvdr paid' field is the amount that CCBH actually paid the provider for the bill.

mh_pymt_amts <- dbGetQuery(conn = db_con, 
                          statement = 
'select 
    a.*,
    e.svc_begin_date,
    e.total_charges, 
    e.prvdr_paid
from (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
left join ac_mh.ccbho_member c on a.mci_uniq_id = c.mci_uniq_id
left join ac_mh.ccbho_claims_paid e on e.recipient_nr = c.ma_recipient_num
where
    svc_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(mh_pymt_amts) <- tolower(names(mh_pymt_amts))

# Reformat dates in R date format

mh_pymt_amts <- mh_pymt_amts %>%
  mutate(svc_begin_date = as.Date(substring(as.character(svc_begin_date), 1, 10), '%Y-%m-%d'))

# Save data

save(mh_pymt_amts, file = paste0(filepath, "/Data/Medicaid claims data/mh_pymt_amts.rda"))
```

### Pharmacy claims since 1/1/2022

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

pharmacy_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select  
    a.*,
    e.fact_rx_nbr, 
    e.dispense_date, 
    e.prescribed_date, 
    e.adjudication_date,
    e.mos, 
    e.mop,
    e.dispense_quantity, 
    e.days_supply,
    e.standardized_paid_amt, 
    e.prescription_nbr,
    e.refill_nbr,
    e.ndc_drug_cd_nbr, 
    g.label_name, 
    g.brand_name, 
    g.generic_name,
    g.drug_class,
    g.dosage_form_code_desc,
    replace(g.drug_strength_desc, chr(0), \' \') as drug_strength_desc,
    g.fdb_generic_classes, 
    g.hic2_desc, 
    g.hic3_desc,
    g.generic_ther_class_desc, 
    g.standard_ther_class_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
  inner join AC_HCDM.DIM_MEMBER c on a.mci_uniq_id = c.mci_uniq_id
  inner join AC_HCDM.FACT_RX e on c.member_nbr = e.member_nbr
  left join ac_hcdm.dim_drug g on e.ndc_drug_cd_nbr = g.ndc_drug_cd_nbr
where
  e.dispense_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(pharmacy_claims) <- tolower(names(pharmacy_claims))

# Reformat dates in R date format

pharmacy_claims <- pharmacy_claims %>%
    mutate_at(vars(dispense_date, prescribed_date, adjudication_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(pharmacy_claims, file = paste0(filepath, "/Data/Medicaid claims data/pharmacy_claims.rda"))
```



