---
title: "Medicaid recertification requirements study - Prepare study data"
output: html_document
---

```{r}

# Clear space

rm(list=ls())

# Load packages

library(dhsDW)
library(ROracle)
library(tidyverse)

# Set chunk options

knitr::opts_chunk$set(eval = FALSE)

options(scipen=999)

# Set file path

filepath <- 'C:/Users/K011014/OneDrive - Allegheny County/Medicaid/Effect of Medicaid recertification deadlines project/Data analysis'
```

# Create base study sample

## Medicaid enrollment spells for all-time, among adults age 18 to 64 who were enrolled in MAGI Medicaid at some point since 4/1/2023

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

sample_enrollment_dates_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)

select distinct
  a.ma_receipt_num, 
  a.mci_uniq_id,
  c.elig_from_date, 
  c.elig_to_date
from 
  medicaid_magi_adults_04012023 a
  left join ac_mh.ccbh_hc_eligibility c on c.ma_receipt_num = a.ma_receipt_num')

# Make var names lowercase

names(sample_enrollment_dates_raw) <- tolower(names(sample_enrollment_dates_raw))

#-------------------------------------------------------------------------
# Clean up the data

# Reformat dates as R date format

sample_enrollment_dates <- sample_enrollment_dates_raw %>%
  mutate_at(vars(elig_from_date, elig_to_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d')) %>%
  
# Remove any enrollment spells where the start date is after the end date. These are data quality issues that I think we can safely remove. 
  
  filter(elig_from_date < elig_to_date) %>%
  
# Recode end dates of 6/6/2079 as missing. 
  
  mutate(elig_to_date = if_else(elig_to_date == as.Date('2079-06-06'), NA, elig_to_date)) %>%
  
# Deduplicate the data by enrollment spell
  
  distinct()
  
  # Save data

save(sample_enrollment_dates, file = paste0(filepath, "/Data/Medicaid enrollment data/sample_enrollment_dates.rda"))
```

## Medicaid recertification deadlines between 4/1/2023 and today, among adults age 18 to 64 who were enrolled in MAGI Medicaid at some point since 4/1/2023

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

sample_recert_deadlines_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)
    
select distinct
    a.ma_receipt_num, 
    a.mci_uniq_id,
    c.ma_renewal_date as annual_recert_deadline, 
    c.ma_sar_date as semiannual_recert_deadline
from 
    medicaid_magi_adults_04012023 a 
    inner join ac_mh.pa_client_elig_daily c on c.recipient_nbr = a.ma_receipt_num
where
    c.ma_renewal_date is not null 
    or c.ma_sar_date is not null')

# Make var names lowercase

names(sample_recert_deadlines_raw) <- tolower(names(sample_recert_deadlines_raw))

#-------------------------------------------------------------------------
# Clean up the data

# Convert dates to R date format

sample_recert_deadlines <- sample_recert_deadlines_raw %>%
  mutate_at(vars(semiannual_recert_deadline, annual_recert_deadline), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d')) %>%
  
# Reshape the data to have one row per deadline
  
  gather(variable, recert_deadline, -ma_receipt_num, -mci_uniq_id) %>%
  select(-variable) %>%
  
# Deduplicate the data
  
  filter(!is.na(recert_deadline)) %>%
  distinct(ma_receipt_num, mci_uniq_id, recert_deadline) %>%
  
# Remove recertification deadlines that are after today
  
  filter(recert_deadline < Sys.Date())
    
# Save data

save(sample_recert_deadlines, file = paste0(filepath, "/Data/Medicaid enrollment data/sample_recert_deadlines.rda"))
```

## Medicaid termination dates and reasons between 4/1/2023 and today, among adults age 18 to 64 who were enrolled in MAGI Medicaid at some point since 4/1/2023

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

sample_terminations_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)

select distinct
    a.*,
    c.batch_date as termination_date, 
    c.ma_term_reason_code as termination_reason_code,
    e.ma_term_reason_desc as termination_reason_desc
from 
    medicaid_magi_adults_04012023 a 
    left join ac_mh.pa_client_elig_daily c on c.recipient_nbr = a.ma_receipt_num 
    left join ac_mh.LKP_PA_MA_TERM_REASON e on e.ma_term_reason_code = c.ma_term_reason_code
where 
    c.change_type = \'Terminated Coverage\'')

# Make var names lowercase

names(sample_terminations_raw) <- tolower(names(sample_terminations_raw))

#---------------------------------------------------------------------------
# Clean up data

# Reformat dates as R date format

sample_terminations <- sample_terminations_raw %>%
  mutate(termination_date = as.Date(substring(as.character(termination_date), 1, 10), '%Y-%m-%d'))

# Save data

save(sample_terminations, file = paste0(filepath, "/Data/Medicaid enrollment data/sample_terminations.rda"))
```

# Clinical Connect alerts data

## Client data coverage dates

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

clinical_connect_coverage_dates <- dbGetQuery(conn = db_con, 
                          statement = 
'select 
   a.mci_uniq_id,
   b.first_date,
   change_date as latest_date
from ac_hcdm.cchie_client_master a
inner join
(select a.mci_uniq_id,
 min(a.load_Date) as first_date
from ac_hcdm.cchie_client_master_hist a
group by mci_uniq_id) b
on a.mci_uniq_id = b.mci_uniq_id')

# Make var names lowercase

names(clinical_connect_coverage_dates) <- tolower(names(clinical_connect_coverage_dates))

#------------------------------------------------------------------------------
# Clean the data

# Reformat dates as R date format

clinical_connect_coverage_dates <- clinical_connect_coverage_dates %>%
  mutate_at(vars(first_date, latest_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(clinical_connect_coverage_dates, file = paste0(filepath, "/Data/Clinical Connect health care data/clinical_connect_coverage_dates.rda"))
```

## Hospital event data

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

clinical_connect_events_raw <- dbGetQuery(conn = db_con, 
                          statement = 
'with medicaid_magi_adults_04012023 as
(select distinct 
    a.ma_receipt_num, 
    e.mci_uniq_id
 from 
    ac_mh.ccbh_hc_eligibility a
    left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
where
    a.category = \'MG\'
    and a.program_status in (\'27\', \'71\', \'91\')
    and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
    and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18)

select distinct
    a.*,
    e.encounter_id,
    e.event_type, 
    e.patient_class, 
    e.admission_type, 
    e.event_date_time,
    e.admit_date_time,
    e.discharge_date_time,
    e.admit_source,
    e.admission_reason,
    e.diagnosis_code_1,
    e.diagnosis_description_1, 
    e.diagnosis_code_2,
    e.diagnosis_description_2, 
    e.diagnosis_code_3,
    e.diagnosis_description_3, 
    e.diagnosis_code_4,
    e.diagnosis_description_4, 
    e.diagnosis_code_5,
    e.diagnosis_description_5
from 
    medicaid_magi_adults_04012023 a
    inner join 
        (select distinct
            mci_uniq_id, 
            uniqueid
        from 
            AC_HCDM.CCHIE_CLIENT_MASTER) c on c.mci_uniq_id = a.mci_uniq_id 
    inner join ac_hcdm.cchie_alert e on LPAD(e.roster_patient_id, 10, \'0\') = c.uniqueid')

# Make var names lowercase

names(clinical_connect_events_raw) <- tolower(names(clinical_connect_events_raw))

#------------------------------------------------------------------------------
# Clean the data

# Reformat dates as R date format

clinical_connect_events <- clinical_connect_events_raw %>%
  mutate_at(vars(event_date_time, admit_date_time, discharge_date_time), ~ as.Date(substring(as.character(.), 1, 10), '%Y%m%d'))

# Save data

save(clinical_connect_events, file = paste0(filepath, "/Data/Clinical Connect health care data/clinical_connect_events.rda"))
```

# Medical examiner data

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

medical_examiner_deaths <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct
    a.*, 
    e.caldr_dt,
    c.ME_ALL_ACT_IND,
    c.ME_ACC_OD_ALL_ACT_IND,
    c.ME_ACC_OD_OPD_ACT_IND,
    c.ME_ACC_OD_NON_OPD_ACT_IND,
    c.ME_ACC_OD_UNK_ACT_IND,
    c.ME_HOM_ACT_IND,
    c.ME_SUIC_ACT_IND,
    c.ME_OTH_ACT_IND
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    inner join dw.fact_client_prog_status_month @DW c on c.mci_uniq_id = a.mci_uniq_id 
    left join dw.dim_time @DW e on e.time_key = c.time_key
where
    c.ME_ALL_ACT_IND + c.ME_ACC_OD_ALL_ACT_IND + c.ME_ACC_OD_OPD_ACT_IND + c.ME_ACC_OD_NON_OPD_ACT_IND + 
    c.ME_ACC_OD_UNK_ACT_IND + c.ME_HOM_ACT_IND + c.ME_SUIC_ACT_IND + c.ME_OTH_ACT_IND > 0
    and e.caldr_dt >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(medical_examiner_deaths) <- tolower(names(medical_examiner_deaths))

# Reformat dates to R date format

medical_examiner_deaths <- medical_examiner_deaths %>%
  mutate_at(vars(caldr_dt), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d')) 

# Save data

save(medical_examiner_deaths, file = paste0(filepath, "/Data/Medicaid claims data/medical_examiner_deaths.rda"))
```

# UPMC claims data

```{r, eval = TRUE}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

upmc_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select 
    a.*,
    c.CLAIM_NBR, 
    c.claim_line_nbr, 
    c.service_begin_date,
    c.ICD_PROCEDURE_ID,
    c.icd_procedure_desc, 
    c.cpt4_code, 
    c.cpt4_desc,
    e.icd_diagnosis_id, 
    e.icd_diagnosis_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    inner join ac_upmc.upmc_claim_proc c on c.mci_uniq_id = a.mci_uniq_id
    left join ac_upmc.upmc_claim_dx e on e.master_member_id = c.master_member_id and e.claim_nbr = c.claim_nbr
where
    c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(upmc_claims) <- tolower(names(upmc_claims))

# Reformat dates to R date format

upmc_claims <- upmc_claims %>%
  mutate_at(vars(service_begin_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(upmc_claims, file = paste0(filepath, "/Data/UPMC claims data/upmc_claims.rda"))
```

# Medicaid claims data

## Physical health claims

```{r, eval = TRUE}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

#--------------------------------------------------------------------------
# Non-ER inpatient care

non_er_inpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_INPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Non-Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(non_er_inpatient_claims) <- tolower(names(non_er_inpatient_claims))

# Reformat dates to R date format

non_er_inpatient_claims <- non_er_inpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(non_er_inpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/non_er_inpatient_claims.rda"))

#--------------------------------------------------------------------------
# ER inpatient care

er_inpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_INPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(er_inpatient_claims) <- tolower(names(er_inpatient_claims))

# Reformat dates to R date format

er_inpatient_claims <- er_inpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(er_inpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/er_inpatient_claims.rda"))

#--------------------------------------------------------------------------
# Non-ER outpatient care

non_er_outpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_OUTPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Non-Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(non_er_outpatient_claims) <- tolower(names(non_er_outpatient_claims))

# Reformat dates as R date format

non_er_outpatient_claims <- non_er_outpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(non_er_outpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/non_er_outpatient_claims.rda"))

#--------------------------------------------------------------------------
# ER outpatient care

er_outpatient_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct 
    a.*,
    c.claim_enc_nbr, 
    c.service_begin_date, 
    c.service_end_date,
    g.promise_specialty_desc,
    k.diag_cd, 
    k.diag_cd_desc
from 
    (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
    left join ac_hcdm.dim_member b on b.mci_uniq_id = a.mci_uniq_id
    inner join AC_HCDM.FACT_PHBH_OUTPATIENT c on c.member_nbr = b.member_nbr
    INNER JOIN AC_HCDM.BRIDGE_GENERIC_SERVICE_CAT d ON (d.SVC_CAT_GRP_NBR_SK = c.SVC_CAT_GRP_NBR)
    INNER JOIN AC_HCDM.DIM_GENERIC_SERVICE_CAT e ON (e.SVC_CAT_NBR_SK = d.SVC_CAT_NBR_SK)
    inner join ac_hcdm.dim_dhs_provider g on (g.dhs_provider_nbr_sk = c.billing_dhs_provider_nbr_sk)
    INNER JOIN AC_HCDM.BRIDGE_DIAGNOSIS_CODE h ON h.DIAG_CD_GRP_NBR_SK = c.DIAG_CD_GRP_NBR_SK
    INNER JOIN AC_HCDM.DIM_DIAGNOSIS_CODE k ON k.DIAG_CD_NBR_SK = h.DIAG_CD_NBR_SK
where 
    e.svc_cat_system = \'ER Indicator\'
    and e.svc_cat_desc = \'Emergency Room Service\'
    and c.service_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')')

# Make var names lowercase

names(er_outpatient_claims) <- tolower(names(er_outpatient_claims))

# Reformat dates as R date format

er_outpatient_claims <- er_outpatient_claims %>%
  mutate_at(vars(service_begin_date, service_end_date), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(er_outpatient_claims, file = paste0(filepath, "/Data/Medicaid claims data/er_outpatient_claims.rda"))
```

## Mental health claims

```{r}

db_con <- dhs_dw(database = "ACPRD1", stored_db = TRUE, stored_creds = TRUE)

#---------------------------------------------------------------------------------
# Substance use disorder service claims

sud_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct
    a.*,
    c.claim_number, 
    c.svc_start_dt, 
    c.svc_end_dt, 
    c.payer, 
    c.office, 
    c.service_code, 
    c.service_name, 
    c.cost_ctr_common_desc, 
    c.service_category, 
    c.prvdr_name, 
    c.cnty_tot, 
    c.tot_units, 
    c.diagnosis_desc, 
    c.diagnosis_common_sort
from (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
inner join dare_health.sud_fy1518 @DW c on a.mci_uniq_id = c.mci_uniq_id
where
    c.svc_start_dt >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(sud_claims) <- tolower(names(sud_claims))

# Reformat dates in R date format

sud_claims <- sud_claims %>%
    mutate_at(vars(svc_start_dt, svc_end_dt), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(sud_claims, file = paste0(filepath, "/Data/Medicaid claims data/sud_claims.rda"))

#---------------------------------------------------------------------------------
# Non-SUD behavioral health claims

mh_claims <- dbGetQuery(conn = db_con, 
                          statement = 
'select distinct
    a.*,
    c.service_rendered_row_id, 
    c.svc_start_dt, 
    c.svc_end_dt, 
    c.payer,
    c.office, 
    c.service_code,
    replace(c.service_name, chr(0), \' \') as service_name,
    c.cost_center, 
    c.service_category, 
    c.provider_name,
    c.cnty_tot, 
    c.tot_units, 
    c.diagnosis_code,
    c.diagnosis_desc, 
    c.diagnosis_common_sort, 
    c.diagnosis_cat
from  (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
inner join dare_health.mh_svc_tableau @DW c on a.mci_uniq_id = c.mci_uniq_id
where
    c.svc_start_dt >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(mh_claims) <- tolower(names(mh_claims))

# Reformat dates in R date format

mh_claims <- mh_claims %>%
    mutate_at(vars(svc_start_dt, svc_end_dt), ~ as.Date(substring(as.character(.), 1, 10), '%Y-%m-%d'))

# Save data

save(mh_claims, file = paste0(filepath, "/Data/Medicaid claims data/mh_claims.rda"))

#---------------------------------------------------------------------------------
# Payments from the Allegheny County behavioral health managed care organization to health care providers

# Note that these amounts are not the cost of care to Medicaid. This is just an alternative way of quantifying the volume of a person's health care utilization. The 'total charges' field is the amount that the provider billed to CCBH. The 'prvdr paid' field is the amount that CCBH actually paid the provider for the bill.

mh_pymt_amts <- dbGetQuery(conn = db_con, 
                          statement = 
'select 
    a.*,
    e.svc_begin_date,
    e.total_charges, 
    e.prvdr_paid
from (select distinct 
        a.ma_receipt_num, 
        e.mci_uniq_id
     from 
        ac_mh.ccbh_hc_eligibility a
        left join ac_mh.t_ahci_clients e on e.ma_receipt_num = a.ma_receipt_num
     where
        a.category = \'MG\'
        and a.program_status in (\'27\', \'71\', \'91\')
        and a.elig_to_date >= to_date(\'01-APR-2023\', \'DD-MON-YYYY\')
        and TRUNC(to_date(\'01-Apr-2023\',\'DD-MON-YYYY\') - e.date_of_birth) / 365.25 >= 18) a
left join ac_mh.ccbho_member c on a.mci_uniq_id = c.mci_uniq_id
left join ac_mh.ccbho_claims_paid e on e.recipient_nr = c.ma_recipient_num
where
    svc_begin_date >= to_date(\'01-JAN-2022\', \'DD-MON-YYYY\')') 

# Make var names lowercase

names(mh_pymt_amts) <- tolower(names(mh_pymt_amts))

# Reformat dates in R date format

mh_pymt_amts <- mh_pymt_amts %>%
  mutate(svc_begin_date = as.Date(substring(as.character(svc_begin_date), 1, 10), '%Y-%m-%d'))

# Save data

save(mh_pymt_amts, file = paste0(filepath, "/Data/Medicaid claims data/mh_pymt_amts.rda"))
```

## Pharmacy claims



